The error in Bayesian phylogenetic reconstruction when speciation co-occurs

### 2019-11-27 fig_results.tex
Moved old pics to pics, use new pretty one-mu-per-figure figures
Rename duplicate figure labels
Add 10 replicate figures
Diminishing returns in ESSes, both figure and text (40x longer MCMC for a decreases of 'ESS < 200' of 50%)
Add figures

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{razzo_results.png}
  \caption{
    The inference error distribution 
    for the different biological
    parameter settings. In all cases, $\lambda = 0.2$ and 
    crown age equals 6. 
  }
  \label{fig:results}
\end{figure}

>>>>

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{figure_1a.png}
  \caption{
    The error distribution for a extinction rate $\mu = 0.0$.
  }
  \label{fig:errors_yule}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{figure_1b.png}
  \caption{
    The error distribution for extinction rate $\mu = 0.15$.
  }
  \label{fig:errors_bd}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{figure_2.png}
  \caption{
    The relation between percentage of observable MBness and
    the inference error. Each point corresponds to the error from
    one posterior tree.
  }
  \label{fig:mbness_and_error}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{figure_3.png}
  \caption{
    The relation between the number of taxa and the
    variance of the inference error. 
    Each point corresponds to the variance of one posterior.
    \richel{move to supplementary materials?}
  }
  \label{fig:n_taxa_and_error_variance}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{figure_4.png}
  \caption{
    The relation between the extinction rate and the
    variance of the inference error. 
    \richel{move to supplementary materials?}
  }
  \label{fig:ext_rate_and_error_variance}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{figure_5.png}
  \caption{
    The relation between the nLTT statistic and the difference
    between the HPD of the true and twin trees' errors. 
    Each point corresponds to the difference of 
    the HPD error between true and twin posterior.
    The grey line is the identity line.
  }
  \label{fig:nltt_and_error_difference}
\end{figure}

### 2019-11-27 razzo-hypotheses.tex
Renamed section to 'hypotheses'
Add figures

\section{Hypothesis}

RJCB expects that the inference error is lowest when the true
tree is generated under MBD parameter settings without multiple-births, i.e
$nu = 0 \vee q = 0$. For such parameters settings, the
true is in practice generated by a BD model, which
matches the tree prior used in inference. Without a mismatch between
true tree prior and assumed tree prior, this source of errors will be absent,
where the other two sources of errors (stochasticity in the
simulation of the alignment and the MCMC algorithm) 
will remain the same.

RJCB expects that, on average, the inference error is highest for
parameter settings with a higher percentage of multiple-births,
i.e. $nu > 0 \wedge q > 0$, as for these settings, 
the mismatch between the the speciation model that
generated the true tree (which is profoundly-MB MBD) is biggest with
the tree prior assumed to be generative (which is BD).

RJCB expects that the inference error is higher for true trees
with a higher percentage of species generated in an MB event,
regardless of the MBD parameters.
These multiple-born species are one of the three (and the most 
interesting) sources of error, as a BD model -as a feature- will never
infer a co-occuring speciation event.

RJCB expects that an increased extinction rate
has a neutral effect on the error made, 
for runs with an equal percentage of MBness,
because extinction affects all species equally.

RJCB expects that an increased number of taxa
has a negative effect on the variance of the errors made, as there
is more information available to base inference on.

RJCB expects that an increased extinction rate
increases the variance of the errors made,
due to the decrease of the number of taxa, reducing the amount of information
to base inference on.

RJCB expects that the nLTT statistic between a true and twin
tree at the start of the pipeline,
will correlate strongly with the difference
between the higest posterior density of the 
errors of the true and twin tree 
generated at the end of the pipeline. 
This hypothesis
stems from the idea of assuming a setup without noise; that is, with
a DNA alignment of infinite length. From such a DNA alignment 
with infinite information, the MCMC
is able to infer the close-to-correct phylogeny.
\richel{if this is true, we can use this shortcut (instead of running
multiple pirouette replicates) to draw stronger conclusions}

@@

>>>>

\section{Hypotheses}

$\mathcal{H}_1$: RJCB expects that the inference error is lowest when the true
tree is generated under MBD parameter settings without multiple-births, i.e
$nu = 0 \vee q = 0$. For such parameters settings, the
true is in practice generated by a BD model, which
matches the tree prior used in inference. Without a mismatch between
true tree prior and assumed tree prior, this source of errors will be absent,
where the other two sources of errors (stochasticity in the
simulation of the alignment and the MCMC algorithm) 
will remain the same. \richel{figure 1a and b}

$\mathcal{H}_2$: RJCB expects that, on average, the inference error is highest for
parameter settings with a higher percentage of multiple-births,
i.e. $nu > 0 \wedge q > 0$, as for these settings, 
the mismatch between the the speciation model that
generated the true tree (which is profoundly-MB MBD) is biggest with
the tree prior assumed to be generative (which is BD). \richel{figure 1a and b}

$\mathcal{H}_3$: RJCB expects that the inference error is higher for true trees
with a higher percentage of species observably generated in an MB event,
regardless of the MBD parameters.
These multiple-born species are one of the three (and the most 
interesting) sources of error, as a BD model -as a feature- will never
infer a co-occuring speciation event. \richel{figure 2}

$\mathcal{H}_4$: RJCB expects that, on average, the inference error is equal for
parameter settings with an equal $\lambda$, $\nu$ and $q$, i.e. extinction
has no effect on the error made, because extinction affects all species equally. 
\richel{figure 1a and b, or a new figure}

$\mathcal{H}_5$: RJCB expects that an increased number of taxa
has a negative effect on the variance of the errors made, as there
is more information available to base inference on.
\richel{figure 3}

$\mathcal{H}_6$: RJCB expects that an increased extinction rate
increases the variance of the errors made,
due to the decrease of the number of taxa, reducing the amount of information
to base inference on.
\richel{figure 4}

$\mathcal{H}_7$: RJCB expects that the nLTT statistic between a true and twin
tree at the start of the pipeline,
will correlate strongly with the difference
between the higest posterior density (HPD) of the 
errors of the true and twin tree 
generated at the end of the pipeline. 
This hypothesis
stems from the idea of assuming a setup without noise; that is, with
a DNA alignment of infinite length. From such a DNA alignment 
with infinite information, the MCMC
is able to infer the close-to-correct phylogeny.
\richel{if this is true, we can use this shortcut (instead of running
multiple pirouette replicates) to draw stronger conclusions}
\richel{figure 5}

@@

$\mathcal{H}_8$: RJCB expects that the JC69 and strict clock 
candidate models, the close

### 2019-11-27 razzo-methods.tex
Add figures

We proceed in the following way: we build simulated datasets generated under the multiple birth model. Then we run a \verb;pirouette; analysis, which will lead to error distributions between the inference posterior and the original simulated trees. Importantly, the \verb;pirouette; analysis includes also a 'twin' parallel pipeline, which will provide a measure of the baseline error due to pure stochasticity, unavoidably occurring in the process.

@@

We simulate the speciation process in continuous time 
using the \verb;mbd_sim; function from the \verb;mbd; R package (\citep{mbd}).
We let parameters vary using all possible combinations of values 
as shown in Table~\ref{tab:simulation_parameters}.
For each parameter setting, 
we generate $2$ \richel{Will need to calculate how many replicates we can do} 
independent reconstructed trees of the same crown age.

>>>>

We proceed in the following way:
under a range a parameter setting, 
we build simulated datasets generated under the multiple birth model. 
Then we run a \verb;pirouette; analysis, 
which will lead to error distributions
between the inference posterior and the original simulated trees. 
Importantly, the \verb;pirouette; analysis includes 
also a 'twin' parallel pipeline, 
which will provide a measure of the baseline error 
due to pure stochasticity, unavoidably occurring in the process.
We then analyse the complete and baseline errors.

@@

We simulate the speciation process in continuous time 
using the \verb;mbd_sim; function from the \verb;mbd; R package (\citep{mbd}).
We let parameters vary using all possible combinations of values 
as shown in Table~\ref{tab:simulation_parameters}.
For each parameter setting, 
we generate [as much as time permits] 
\richel{Will need to calculate how many replicates we can do} 
independent reconstructed trees of the same crown age.

### 2019-11-27 razzo-results.tex
Add figures

The inference error made for each of the parameter combinations
is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}. 
For both extinction rates,
we find that the error increases as $\nu$
and $q$ increase. 
Also in line with RJCB's predictions, 
we find no difference between the two extinction rates.

>>>>

The inference error made for each of the parameter combinations
is shown in Figures ~\ref{fig:errors_yule} and ~\ref{fig:errors_bd}.

\richel{if true:}The lowest errors are at settings where $nu = 0 \vee q = 0$,
confirming $\mathcal{H}_1$.

\richel{if true:}For both extinction rates,
we find that the error increases as $\nu$
and $q$ increase, confirming $\mathcal{H}_2$

\richel{if true:}There is no difference between the
errors for the two different extinction rates, which
confirms $\mathcal{H}_4$

[other hypotheses being rejected/accepted here]
